Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-04 paper/main.tex
Updates

As we saw in the previous section, depending on the type of variable, different
conditional measures of association are used. However, no generalized measure
of association exists for mixed data. In this section, we present a novel
measure of partial association for mixed data. This measure of association is
analogous to partial correlation coefficient where a residualization method is
combined with Pearson correlation coefficient. Here, we use a mixed data
residualization method \citep{Ankan2023} and combine it with a multivariate
measure of association based on canonical correlations called Pillai's Trace
{Pillai1955}.

@@

\section{Web Tool}
\label{sec:web}

To allow users to apply this method to their own dataset, we also built an
interactive web browser based tool to allow them to interactively construct
models. The user starts by uploading their dataset which initializes an empty
DAG on the variables in the dataset. Users can then set the desired p-value
and measure of association threshold. The tools then shows a red edge between
all the variables that are correlated in the data but the correlation is not
explained by the current DAG. The thickness of the edge represents the strength
of this correlation, which would help the user in prioritizing edges to add.

The tool also computes the
Shipley's C value at any point.

The edge
weight of the edge represents how high this strength is. 
Based on these shown potential edges, the user can select the edge that
they want to add and decide the direction for it. If the tool detects an
incorrect edge in the graph, it is shown in black. Once the user is happy with
the graph they have they can export it into different formats to use it for
further computation.

\begin{figure}[t!]
	\centering
	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
	\caption{A screenshot of the web tool for constructing the model. Users
		can upload their dataset after which the tool creates an empty
		graph and shows all pair of variables which are associated in
		the model using undirected red edges with the strength of
		association represented using edge width. Users can then add
		edges to the model (shown in green). Unnecessary edges are shown in 
		black.}
\end{figure}

>>>>

In the previous section, we discussed how different types of variables require
different conditional measures of association. However, no unified measure
exists for mixed data. In this section, we introduce a novel measure of partial
association for mixed data, by extending the idea of partial correlation
coefficient to mixed data. Specifically, our approach combines a mixed data
residualization method \citep{Ankan2023} with Pillai's Trace
\citep{Pillai1955}, a multivariate measure of association based on canonical
correlations.

@@

\section{Web Tool}
\label{sec:web}
To enable users to apply this method to their own datasets, we developed an
interactive web-based tool (shown in Fig.~\ref{fig:web}) for constructing
models. Users begin by uploading their dataset, which initializes an empty DAG
with nodes corresponding to the dataset\'s variables. They can then specify a
p-value threshold and a measure of association threshold.

Using the specified thresholds, the tool visually highlights unexplained
correlations by displaying red edges between variables where correlation exists
in the data but the current DAG is not able to explain. The thickness of these
edges represents the strength of the correlation, helping users prioritize
which edges to add. Similarly, if an edge in the graph is detected to be
unnecessary, it is highlighted in black. Based on this information, users can
select to remove unnecessary edge or select a potential edge to add and specify
the orientation of the edge. The tool computes Shipley’s C \citep{Shipley2000}
value at each change to determine the overall fit of the model to the data.
Once satisfied with the constructed DAG, users can export the model for further
analysis.

\begin{figure}[t!]
	\centering
	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
	\caption{A screenshot of the web tool for constructing the model. Users
		can upload their dataset after which the tool creates an empty
		graph and shows all pair of variables which are associated in
		the model using undirected red edges with the strength of
		association represented using edge width. Users can then add
		edges to the model (shown in green). Unnecessary edges are shown in 
		black.}
	\label{fig:web}
\end{figure}
